### 2015-01-22 ASP.tex
correction1

This simple language has many advantages that make it one of the most effective in terms of knowledge representation, reasoning and declarative problem solving. 
\annotMF{Source ? Sinon : “one of the most” $\Rightarrow$ “very”}

@@

\annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}

The other classes remain constant as they meet some informal agreement. In general, variables start with a capital letter and contain letters and numbers, but often only one capital letter is used (\eg X, Y, ...).
The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter (\eg f, g, a, b, ...).
String surrounded by quotation marks are also considered as constants (\eg "f", "g", "a b c", ...).

\annotMF{Je ne suis pas sûr que l'explication sur les prédicats soit pertinente. Si on la garde, il faudrait la reformuler.}
What seems a little fuzzy is the concept of predicate. Indeed the word predicate can be an innovation of the new grammar. Let's consider the following sentence: 
\begin{tabbing}
 \textit{Socrates was an athenian.} 
\end{tabbing}
\textit{Socrate} is the subject and \textit{was an athenian} is the predicat. This sentence can be noted in ASP by :
\begin{tabbing}
 \textit{athenian(Socrate) }$\leftarrow ~ \top$. \hspace{2cm} with $\top$ is the symbol of True.
\end{tabbing}

\item \textbf{The rules}
As previously mentioned, a rule has the following form:
%\begin{equation} 
%head ~ \leftarrow ~ body.
%\tag{\eqref{eq1ASP}}
%\end{equation}
\begin{equation} \label{eq3ASP}
A ~ \leftarrow ~ B^+, ~ \ASPnot B^-.
\end{equation}
where the part on the left of the arrow is called $head$
and the right part is called $body$.
This rule is then equivalent to:
\begin{equation} \label{eq2ASP}
L_{0} \leftarrow L_{1},...,L_{m}, \ASPnot L_{m+1},..., \ASPnot L_{n}.
\end{equation}
And this rule states:
\[
\left.
    \begin{array}{ll}
        \text{If} ~ L_{1},...,L_{m} ~ \text{are \textbf{true}}\\
        \text{and} ~ L_{m+1},...,L_{n} ~ \text{are \textbf{false}}
    \end{array}
\right \} \text{Then} ~ L_{0} ~ \text{is } \textbf{true} 
\]
with $L_{i}$ literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
Therefore, $head$ is also called the \emph{conclusion} of the rule,
and $body$ is its \emph{premise}.

There are some special cases of rules \cite{Vladimir,baral2003knowledge}:
\begin{itemize}
\item A rule is said \textbf{constant} or \emph{ground} if all of its literals are constants;
\item A \textbf{fact} or a \textbf{reality} is a rule with an empty body.
  It can be written without the central arrow ($\leftarrow$) or with a body equal to $\top$:
\begin{equation} 
L_{0} \leftarrow \top. \qquad \text{or} \qquad L_{0}.
 \label{eq5ASP}
\end{equation}

\item A \textbf{constraint} is a rule where the head equals false ($\bot$).
  In this case, the head is often not represented and the constraint will be written generally as:
\begin{equation} 
  \leftarrow L_{1},...,L_{m}, \ASPnot L_{m+1},..., \ASPnot L_{n}.
  \label{eq6ASP}
\end{equation}

We say that a set of literals $X$ violates the constraint \eqref{eq6ASP} if $\{L_{1},...,L_{m}\} \subseteq X$ and $\{L_{m+1},...,L_{n}\} \nsubseteq X$.
If a program $\pi$ contains such a constraint $r$, then $X$ is an answer set of $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$ and $X$ does not violate $r$.

\item A \emph{cardinality constraint} is a rule of the following form:
\begin{equation} 
 ~ l ~\{q_{1},~ ... ,~ q_{m}\}~ u \leftarrow body.
 \label{eq7ASP}
\end{equation}
with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$).
The literals $q_i$ can also appear in the body of the rule or in other rules,
and $l$ and $u$ are the lower and upper bounds of the number of these literals
that can appear in the answer sets.
%cardinality of the subsets of $ \{q_{1}, ... , q_{m}\}$ that satisfy the answer sets.
Therefore, a cardinality constraint is satisfied in an answer set $X$, if the number of atoms of $\{q_{1}, ... , q_{m}\}$ belonging to $X$ is between $l$ and $u$,
or, in other words, if:
\begin{equation}
  l \leq | \{q_{1},... , q_{m}\} \cap X | \leq u
\end{equation}
where $\cap$ is the symbol of the intersection of sets
and $|A|$ is the cardinality of the set $A$.
\modMF{%
In the following, if they are not explicitly given,
we consider that $l$ defaults to $0$ and $u$ defaults to $\infty$.
\annotMF{Les cardinalités ne sont pas seulement utilisées de cette manière ! Dans notre code, on les trouve dans la tête et dans le corps des règles.}
\todo{À compléter : cardinalités dans le corps des règles}
}%
\end{itemize}

@@

We can consider that the construction of models is one of the fundamental components of the scientific process. It concerns all systems we seek to control. A model has two main features \cite{Glimpse}:
\begin{itemize}
\item it is a simplification of a given system,
\item it allows an action on the system.
\annotMF{Quel genre d'action ?}
\end{itemize}
Models offer the possibility to provide a solution to a problem identified as such.
\reformuler{This concept model can address another angle issues related to representation process.}

@@

\item \textbf{Solving an ASP program}
\modMF{%
The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
that usually work together:
}%
the grounder is used to remove variables in order to achieve an equivalent but constant program,
and the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
Amongst all available grounders for ASP, we can name
\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
and the solvers include
\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...

\reformuler{
The combined use of the grounder and the solver specifies major programs in a compact, using rules with schematic variables and other abbreviations. Both systems employ grounding algorithms that work quickly and simplify the program.
}
For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

>>>>

This simple language has many advantages that make it very effective in terms of knowledge representation, reasoning and declarative problem solving. 
\annotMF{Source ? Sinon : “one of the most” $\Rightarrow$ “very”} \repEB{OK}

@@

\annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
\repEB{T’as raison au début j’ai écrit tout le code en mode $math$ et puis j’ai changé juste la partie du code et j’ai oublié de faire la mise à jour de cette partie de définition d’ASP}
The other classes remain constant as they meet some informal agreement. In general, variables start with a capital letter and contain letters and numbers, but often only one capital letter is used (\eg X, Y, ...).
The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter (\eg f, g, a, b, ...).
String surrounded by quotation marks are also considered as constants (\eg "f", "g", "a b c", ...).

\annotMF{Je ne suis pas sûr que l'explication sur les prédicats soit pertinente. Si on la garde, il faudrait la reformuler.}
What seems a little fuzzy is the concept of predicate. Indeed the word predicate can be an innovation of the new grammar. Let's consider the following sentence: 
\begin{tabbing}
 \textit{Socrates was an athenian.} 
\end{tabbing}
\textit{Socrate} is the subject and \textit{was an athenian} is the predicat. This sentence can be noted in ASP by :
\begin{tabbing}
 \texttt{athenian(Socrate) $\leftarrow ~ \top$. }\hspace{2cm} with \texttt{$\top$}  is the symbol of True.
\end{tabbing}

\item \textbf{The rules}
As previously mentioned, a rule has the following form:
\texttt{
\begin{equation*} \label{eq3ASP}
 A ~ \leftarrow ~ B^+, ~ \ASPnot B^-.
\end{equation*}
}

where the part on the left of the arrow is called $ head$
and the right part is called $body$.
This rule is then equivalent to:
\texttt{
\begin{equation*} \label{eq2ASP}
l_{0} \leftarrow l_{1},...,l_{m}, \ASPnot l_{m+1},..., \ASPnot l_{n}.
\end{equation*}
}
And this rule states:
\[
\left.
    \begin{array}{ll}
        \text{If} ~ \texttt{$l_{1},...,l_{m}$}  ~ \text{are \textbf{true}}\\
        \text{and} ~ \texttt{$l_{m+1},...,l_{n}$} ~ \text{are \textbf{false}}
    \end{array}
\right \} \text{Then} ~ \texttt{$ l_{0} $} ~ \text{is } \textbf{true} 
\]
with \texttt{$l_{i}$} literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
Therefore, $head$ is also called the \emph{conclusion} of the rule,
and \texttt{body} is its \emph{premise}.

There are some special cases of rules \cite{Vladimir,baral2003knowledge}:
\begin{itemize}
\item A rule is said \textbf{constant} or \emph{ground} if all of its literals are constants;
\item A \textbf{fact} or a \textbf{reality} is a rule with an empty body.
  It can be written without the central arrow ($\leftarrow$) or with a body equal to $\top$:
  \texttt{
\begin{equation} 
l_{0} \leftarrow \top. \qquad \text{or} \qquad l_{0}.
 \label{eq5ASP}
\end{equation}
}

\item A \textbf{constraint} is a rule where the head equals false ($\bot$).
  In this case, the head is often not represented and the constraint will be written generally as:
\begin{equation} 
  \leftarrow l_{1},...,l_{m}, \ASPnot l_{m+1},..., \ASPnot l_{n}.
  \label{eq6ASP}
\end{equation}

We say that a set of literals $X$ violates the constraint \eqref{eq6ASP} if $\{l_{1},...,l_{m}\} \subseteq X$ and $\{l_{m+1},...,l_{n}\} \nsubseteq X$.
If a program $\pi$ contains such a constraint $r$, then $X$ is an answer set of $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$ and $X$ does not violate $r$.

\item A \emph{cardinality constraint} is a rule of the following form:
\begin{equation} 
 ~ l ~\{q_{1},~ ... ,~ q_{m}\}~ u \leftarrow body.
 \label{eq7ASP}
\end{equation}
with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$). \repEB{ This form of rule \ref{eq7ASP} is used when the body can lead to several $q_i$ atoms and also when the programmer need to have more than an atom of type $q_i$. But that the number of this atoms $q_i$ should be between $l$ and $m$. That means that it is possible to find the same atom repeated many times at one answer set.}  \\
In addition the literals $q_i$ can also appear in the body of the rule,
and $l$ and $u$ are the lower and upper bounds of the number of these literals
that can appear in the answer sets.
%cardinality of the subsets of $ \{q_{1}, ... , q_{m}\}$ that satisfy the answer sets.
Therefore, a cardinality constraint is satisfied in an answer set $X$, if the number of atoms of $\{q_{1}, ... , q_{m}\}$ belonging to $X$ is between $l$ and $u$,
or, in other words, if:
\begin{equation}
  l \leq | \{q_{1},... , q_{m}\} \cap X | \leq u
\end{equation}
where $\cap$ is the symbol of the intersection of sets
and $|A|$ is the cardinality of the set $A$.
\modMF{%
In the following, if they are not explicitly given,
we consider that $l$ defaults to $0$ and $u$ defaults to $\infty$.
\annotMF{Les cardinalités ne sont pas seulement utilisées de cette manière ! Dans notre code, on les trouve dans la tête et dans le corps des règles.}
\todo{À compléter : cardinalités dans le corps des règles}
}%
\end{itemize}

@@

We can consider that the construction of models is one of the fundamental components of the scientific process. It concerns all systems we seek to control. A model has two main features \cite{Glimpse}:
\begin{itemize}
\item it is a simplification of a given system,
\item it allows an action on the system \repEB{like resolving problems, searching for a path, identifying states ... }
\annotMF{Quel genre d'action ?} 
\end{itemize}
This concept model can address another angle issues related to representation process.

@@

\item \textbf{Solving an ASP program}
\modMF{%
The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
that usually work together:
}%
first the grounder is used to remove variables in order to achieve an equivalent but constant program,
then the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
Amongst all available grounders for ASP, we can name
\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
and the solvers include
\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...

\reformuler{
The combined use of the grounder and the solver specifies major programs in a compact, using rules with schematic variables and other abbreviations. Both systems employ grounding algorithms that work quickly and simplify the program.
}
\repEB{c'est à ajouter dans le paragraphe ou à reformuler?}
For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

### 2015-01-22 PH.tex
correction1

\todo{Glu pour la définition \ref{def:NextState}}

\begin{definition}[Next states]
\label{def:NextState}
Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
one of its states. The set of the next possible states for $s$ are computed as follows:
\begin{center}
$next(s) = \{s[b_k /b_j ]| \exists (a_i , b_j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa\}$
\end{center}
\todo{Définir la notation $s[b_k /b_j ]$}
\annotMF{Est-ce que $next(s)$ est vraiment utile ? On ne s'en ressert pas après.}
\end{definition}

@@

\begin{definition}[Stable state or fix point]
\label{def:FixPoint1}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$s \in \PHl$ be state;
$s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
\annotMF{“Stable state” ? “Fix point” ? “Fixed point” ?}
\end{definition}

@@

\begin{definition}[Reachability question]
\label{def:reachability}

 If $s \in \PHl$ is a state and $A \subseteq \Proc$ a set of processes,
 we note $\mathcal{P}(s, A)$ the \emph{the reachability question}:
 \begin{center}
  $\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$
 \end{center}
\todo{Il faut définir $\Proc$}
\annotMF{Il serait peut-être mieux de parler de “reachability property” plutôt que “question”.}
\end{definition}

>>>>

\todo{Glu pour la définition \ref{def:NextState}} \repEB{ça veut dire quoi Glu?}

\begin{definition}[Next states]
\label{def:NextState}
Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
one of its states. The set of the next possible states for $s$ are computed as follows:
\begin{center}
$next(s) = \{s[b_k /b_j ]| \exists (a_i , b_j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa\}$
\end{center}
\todo{Définir la notation $s[b_k /b_j ]$}
\annotMF{Est-ce que $next(s)$ est vraiment utile ? On ne s'en ressert pas après.} \repEB{Je crois que si, pour le définition du point fixe et pour comprendre l'évolution du réseau d'un état à un autre}
\end{definition}

@@

\begin{definition}[Stable state or fix point]
\label{def:FixPoint1}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$s \in \PHl$ be state;
$s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
\annotMF{“Stable state” ? “Fix point” ? “Fixed point” ?} \repEB{Je pense garder fix point}
\end{definition}

@@

\begin{definition}[Reachability property]
\label{def:reachability}

 If $s \in \PHl$ is a state and $A \subseteq \Proc$ \repEB{ with $\Proc$ is a set of processes that we find at most one process of each sort $\{a_i, b_j,..., z_k\}$ }
 we note $\mathcal{P}(s, A)$ the \emph{the reachability property}:
 \begin{center}
  $\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$
 \end{center}
\todo{Il faut définir $\Proc$}
\annotMF{Il serait peut-être mieux de parler de “reachability property” plutôt que “question”.}  \repEB{ Je l'ai changé} 
\end{definition}

### 2015-01-22 comparaison.tex
correction1

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following methods, that are detailed below:
GINsim (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
LIBDDD (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
Rocca's et al. method \cite{roccaasp}
and PINT \cite{PMR12-MSCS}.
\annotMF{Il y a une confusion entre méthodes et outils ici je pense.}
\annotMF{Il faut trouver un autre nom pour “Rocca's method”.}
Each method use a specific kind of representation:
logical regulatory network for GINsim,
Instantiable Transition Systems for LIBDDD,
state transition network for Rocca's et al. method
and PH for Pint and our method.

\modMF{%
For this comparative study, we focus biological network of different sizes:
a tadpole tail resorption (TTR) model with 12 biological components~\cite{khalis2009smbionet},
an ERBB receptor-regulated G1/S transition (ERBB) model with 20 components~\cite{Samaga2009}
and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.
}%
These models were chosen to be of different sizes:
from small (12 components) to large (40 components).
We note however that the PH models considered may contain more sorts than
the original number of biological components, due to the use of
“cooperative sorts”, which allow to model Booelan gates but do not necessarily
have a biological meaning.
The different model representations required have been obtained by translations
from the PH
ensuring the conservation of the dynamical properties.
All these translations are available with Pint.
\todo{URL vers Pint et les autres outils}
All results alongside with more detailed specifications of the models
are given in table \ref{tab:reachability}.
The methods and the results provided by each of them are detailed in the following.
The overall results shows that our method is efficient in computing reachability
from a given initial state;
furthermore, it sometimes provides more informations that the other existing ones.

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l||c|c|c||c|c|c|c|c|}
  \hline
   Model& \#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
  \hline
  TTR & 42 & $2^{19}$ & state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
  \hline
  ERBB & 152 & $2^{70}$ & state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
  \hline
  ERBB & 152 & $2^{70}$ & process & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
  \hline
  TCR & 156 & $2^{73}$ & process & - & 0m0.014s & out & - & 0m05.02s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

\todo{Rajouter le nombre de sortes dans le tableau ou la description}
\annotMF{Il faut expliciter ce qu'on entend par “state” et “process”. Il fudrait peut-être aussi changer les termes ; je propose : “full state” et “partial state”.}

\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau}

@@

\item \textbf{\textsc{GINsim}} is a software for the edition, simulation and analysis
of gene interaction networks.
It allows to compute all reachable stable states instantly;
however, it is not possible to compute all stable states independently from the initial state.
Regarding the reachability problem, \textsc{GINSim} only allows to check
full states, because of its approach which consists in computing
all the state-transition graph
\todo{Seulement la partie accessible ou l'ensemble du graphe ?}
and then check if there exists a path between two given states.
Therefore, it was not possible to perform reachability checks on partial states.
Small state-transition graphs can also be displayed by this tool.

@@

We developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
\modMF{%
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
}%
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop.
\annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
\begin{lstlisting}[numbers=none]
iclingo find\_path.lp <network\_name>.lp --imax=n
\end{lstlisting}
\annotMF{Cette ligne de commande ne me semble pas nécessaire. Mais si on la garde, il faut en donner un autre exemple plus tôt, sans l'argument -{}-imax=n.}
Some interesting values can be given to this integer \texttt{n},
which sets the maximum number of steps.
For example, the total number of states is a maximum which is obvious,
as it will never be exceeded by a minimum path,
but too hight to be very interesting.
The total number of sorts can also be an interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks,
or, by using the same reasoning, the total number of processes.
\modMF{%
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.
}%

>>>>

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al. \cite{roccaasp} which was developed ALSO on ASP.
\annotMF{Il y a une confusion entre méthodes et outils ici je pense.} \repEB{J'ai modifié}
\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
Each method use a specific kind of representation:
logical regulatory network for GINsim,
Instantiable Transition Systems for LIBDDD,
state transition network for Rocca's et al. method
and Process Hitting (PH) for Pint as well as for our method.

\modMF{%
For this comparative study, we focus biological network of different sizes:
a tadpole tail resorption (TTR) model with 12 biological components~\cite{khalis2009smbionet},
an ERBB receptor-regulated G1/S transition (ERBB) model with 20 components~\cite{Samaga2009}
and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.
}%
These models were chosen to be of different sizes:
from small (12 components) to large (40 components).
We note however that the PH models considered may contain more sorts than
the original number of biological components, due to the use of
“cooperative sorts”, which allow to model Booelan gates but do not necessarily
have a biological meaning.
The different model representations required have been obtained by translations
from the PH
ensuring the conservation of the dynamical properties.
All these translations are available with \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} }
\todo{URL vers Pint et les autres outils}
All results alongside with more detailed specifications of the models
are given in table \ref{tab:reachability}.
The methods and the results provided by each of them are detailed in the following.
The overall results shows that our method is efficient in computing reachability
from a given initial state;
furthermore, it sometimes provides more informations that the other existing ones.

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

\todo{Rajouter le nombre de sortes dans le tableau ou la description} \repEB{avec ou sans les sortes coopératives? car pour PH y'en a plus de composants que pour les autres modèles}
\annotMF{Il faut expliciter ce qu'on entend par “state” et “process”. Il fudrait peut-être aussi changer les termes ; je propose : “full state” et “partial state”.} \modifEB{Je préfère ta proposition mais tu vois ça prend plus d'espace et le tableau dépasse les limites }

\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau} \repEB{En fait pour la méthode de Rocca le script qu'il m'ont envoyé génére quelques erreurs lors de la compilation. pour libddd y'avait pas l'option (--fixpoint) ne génère rien comme résultat et GINsim répond très rapidement mais il faudrait calculer tout le graphe de transition avant. }

@@

\item \textbf{\textsc{GINsim}} is a software for the edition, simulation and analysis
of gene interaction networks.
It allows to compute all reachable stable states instantly;
however, it is not possible to compute all stable states independently from the initial state.
Regarding the reachability problem, \textsc{GINSim} only allows to check
full states, because of its approach which consists in computing
all the state-transition graph
\todo{Seulement la partie accessible ou l'ensemble du graphe ?} \repEB{Je pense que oui mais je n'arrive pas à le m'assurer car il m'a fallut les états initiaux pour tester l'atteignabilité à ces états stables}
and then check if there exists a path between two given states.
Therefore, it was not possible to perform reachability checks on partial states.
Small state-transition graphs can also be displayed by this tool.

@@

We developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
\modMF{%
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
}%
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop. This is possible with the option \texttt{"--imax=n"}.
\annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
\annotMF{Cette ligne de commande ne me semble pas nécessaire. Mais si on la garde, il faut en donner un autre exemple plus tôt, sans l'argument -{}-imax=n.} \modifEB{ok t'as raison je l'ai enlevée et remplacée par juste l'option \texttt{"--imax=n"}}
Some interesting values can be given to this integer \texttt{n},
which sets the maximum number of steps.
For example, the total number of states is a maximum which is obvious,
as it will never be exceeded by a minimum path,
but too hight to be very interesting.
The total number of sorts can also be an interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks,
or, by using the same reasoning, the total number of processes.
\modMF{%
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.
}%

### 2015-01-22 intro.tex
correction1

The objectives of the work presented in this paper are the following.
Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fixed points \cite{wuensche1998genomic} \annotMF{Est-ce que cette ref est pertinente ? Elle me semble plutôt axée sur la notion d'attracteur}).
For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fixed points.
The second phase of our work consists in computing the dynamics. It consists in determining from a known initial state the possible next states of the PH model. Then we verify if we can reach a specific state for one or several component (genes or proteins). The results are ensured to respect strictly the PH dynamics.

@@

\modMF{%
Section \ref{defs} defines the paradigm of Answer Set Programming,
and the Process Hitting framework.
We then use ASP in section \ref{fixpoint} to tackle the enumeration
of all fixed points in a model,
and in section \ref{dynamics} in order to compute the dynamics of a model
and check reachability properties.
Finally, a comparison of the methods developed in this paper is
proposed in section \ref{comparison},
and in section \ref{ccl} we conclude and talk about future work.
}%

>>>>

The objectives of the work presented in this paper are the following.
Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fix points \cite{wuensche1998genomic} \annotMF{Est-ce que cette ref est pertinente ? Elle me semble plutôt axée sur la notion d'attracteur}). \modifEB{Je l’ai mise car en effet je pense que le point fixe est un cas spécial des attracteurs.}
For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fix points.
The second phase of our work consists in computing the dynamics. It consists in determining from a known initial state the possible next states of the PH model. Then we verify if we can reach a specific state for one or several component (genes or proteins). The results are ensured to respect strictly the PH dynamics.

@@

\modMF{%
Section \ref{defs} defines the paradigm of Answer Set Programming,
and the Process Hitting framework.
We then use ASP in section \ref{fixpoint} to tackle the enumeration
of all fix points in a model,
and in section \ref{dynamics} in order to compute the dynamics of a model
and check reachability properties.
Finally, a comparison of the methods developed in this paper is
proposed in section \ref{comparison},
and in section \ref{ccl} we conclude and talk about future work.
}%
